---
knit: "bookdown::render_book"
title: "R Code for Mastering 'Metrics"
author: ["Jeffrey B. Arnold"]
description: >
  R code to reproduce analyses in Mastering Metrics 
  by Angrist and Pischke.
site: bookdown::bookdown_site
bibliography: ["masteringmetrics.bib"]
github-repo: "jrnold/r4ds-exercise-solutions"
url: 'http\://jrnold.github.io/r4ds-exercise-solutions'
twitter-handle: jrnld
documentclass: book
editor_options: 
  chunk_output_type: console
---

# Welcome {-}

This work contains R code to reproduce many of the analyses in *Mastering 'Metrics* by Joshua D. Angrist and Jörn-Steffen Pischke [@AngristPischke2014].
This work provides R translations of the replication code available at [masteringmetrics.com](http://masteringmetrics.com/resources/).

## Install {-}

To install all packages used in the examples in this work
and the datasets from *Mastering 'Metrics* run
```{r eval=FALSE}
devtools::install_github("jrnold/masteringmetrics", subdir = "masteringmetrics")
```

## License {-}

The text of this work is licensed under the [Creative Commons Attribution 4.0 International License](http://creativecommons.org/licenses/by/4.0/).
The R Code in this work is licensed under the [MIT License](https://opensource.org/licenses/MIT).

## Colonophon {-}

The book is powered by <https://bookdown.org> which makes it easy to turn R markdown files into HTML, PDF, and EPUB.

This book was built with:

```{r}
devtools::session_info(c("tidyverse"))
```

```{r include=FALSE, cache=FALSE}
```
```{r include=FALSE, cache=FALSE}
```
