# Minneapolis Domestic Violence Experiment

This replicates Table 3.3 of *Mastering 'Metrics*, which replicates the Minneapolis Domestic Violence Experiment [@ShermanBerk1984,@Angrist2006].

Load necessary packages.
```{r library,message=FALSE}
library("tidyverse")
```

Load the MDVE data.
```{r mdve}
data("mdve", package = "masteringmetrics")
```

Randomized assignments (i.e. what are police assigned to do) are in the `assigned` column.
Actual outcomes (i.e. what action do the police actually take) is in the `outcome` column.
gen outcome = "Arrest" if T_FINAL == 1
replace outcome = "Advise" if T_FINAL == 2
replace outcome = "Separate" if T_FINAL == 3
replace outcome = "Other" if T_FINAL == 4
gen total = 1
```{r}
mdve <- mutate(mdve,
               assigned = case_when(
      T_RANDOM == 1 ~ "Arrest",
      T_RANDOM == 2 ~ "Advise",
      T_RANDOM == 3 ~ "Separate"
    ),
      outcome = case_when(
        T_FINAL == 1 ~ "Arrest",
        T_FINAL == 2 ~ "Advise",
        T_FINAL == 3 ~ "Separate",
        T_FINAL == 4 ~ "Other"
      ),
      coddled_a = assigned != "Arrest",
      coddled_o = outcome != "Arrest"
    ) %>%
  filter(outcome != "Other")
```

Assigned and delivered treatments in the MDVE:
```{r}
mdve_summary <-
  mdve %>%
  count(assigned, outcome) %>%
  group_by(assigned) %>%
  mutate(p = n / sum(n))
print(mdve_summary, n = nrow(mdve_summary))
```

Assigned proportions in the MDVE:
```{r}
mdve_assigned <- mdve %>%
  count(assigned) %>%
  mutate(p = n / sum(n))
mdve_assigned
```

Delivered treatments in the MDVE:
```{r}
mdve_outcome <- mdve %>%
  count(outcome) %>%
  mutate(p = n / sum(n))
mdve_outcome
```

Probability of being coddled, given being assigned the coddled treatment:
```{r}
mdve_coddled <- mdve %>%
  count(coddled_a, coddled_o) %>%
  group_by(coddled_a) %>%
  mutate(p = n / sum(n))
mdve_coddled
```

IV first stage,
$$
E[D_i | Z_i = 1] - E[D_i | Z_i = 0] .
$$
```{r}
filter(mdve_coddled, coddled_o, coddled_a)$p -
  filter(mdve_coddled, coddled_o, !coddled_a)$p
```

The response variable is not provided, so the full 2SLS is not estimated here.

## References

-   <http://masteringmetrics.com/wp-content/uploads/2015/02/MDVE_Table33.do>
-   <http://masteringmetrics.com/wp-content/uploads/2015/02/ReadMe_MDVE.txt>
```{r include=FALSE, cache=FALSE}
```
