---
output: html_document
editor_options: 
  chunk_output_type: console
---

# MLDA Difference-in-Difference

Difference-in-difference estimates of the effect of the minimum legal drinking age (MLDA) on mortality [@DuMouchelWilliamsZador1987,@NorbergBierutGrucza2009].
This replicates the analyses in Tables 5.2 and 5.3 in *Mastering 'Metrics*.

Load necessary libraries.
```{r setup,message=FALSE}
library("tidyverse")
library("haven")
library("rlang")
library("broom")
```

```{r deaths}
data("deaths", package = "masteringmeterics")
```


## Table 5.2

Regression DD Estimates of MLDA-Induced Deaths among 18-20 Year Olds, from 1970-1983

```{r}
dtypes <- c("all", "MVA", "suicide", "homicide")
```


```{r}
run_reg <- function(i) {
  data <- filter(deaths, year <= 1983, agegr == "18-20 yrs", dtype == i)
  mods <- list(
    "No trends, no weights" = 
      lm(mrate ~ legal + state + year, data = mutate(data, year = factor(year))),
    "Time trends, no weights" = 
      lm(mrate ~ legal + state * year, data = data),
    "No trends, weights" =
      lm(mrate ~ legal + state + year, data = mutate(data, year = factor(year)),
         weights = data$pop),
    "Time trends, weights" = 
      lm(mrate ~ legal + state * year, data = data, weights = data$pop)
  )
  map_df(mods, tidy) %>%
    filter(term == "legal") %>%
    select(estimate, std.error) %>%
    mutate(response = i,
           model = names(mods))
}
                 
```

**TODO: need to add clustering by state**

```{r}
mlda_dd <- map_df(dtypes, run_reg)
```


## Table 5.3

Regression DD Estimates of MLDA-Induced Deaths among 18-20 Year Olds, from 1970-1983, controlling for Beer Taxes

```{r}
run_reg <- function(i) {
  data <- filter(deaths, year <= 1983, agegr == "18-20 yrs", dtype == i)

  mods <- list(
    "No time trends" = 
      lm(mrate ~ legal + beertax + state + year, data = mutate(data, year = factor(year))),
    "Time trends" = 
      lm(mrate ~ legal + beertax + state * year, data = data),
  )
  map_df(mods, tidy) %>%
    filter(term == "legal") %>%
    select(estimate, std.error) %>%
    mutate(response = i,
           model = names(mods))
}
                 
```

## References

- <http://masteringmetrics.com/wp-content/uploads/2015/01/analysis.do>
- <http://masteringmetrics.com/wp-content/uploads/2015/01/ReadMe_MLDA_DD.txt>
