---
output: html_document
editor_options: 
  chunk_output_type: console
---

# MLDA Difference-in-Difference

Difference-in-difference estimates of the effect of the minimum legal drinking age (MLDA) on mortality [@DuMouchelWilliamsZador1987;@NorbergBierutGrucza2009].
This replicates the analyses in Tables 5.2 and 5.3 in *Mastering 'Metrics*.

Load necessary libraries.
```{r setup,message=FALSE}
library("tidyverse")
library("haven")
library("rlang")
library("broom")
library("clubSandwich")
```

```{r deaths}
data("deaths", package = "masteringmetrics")
```
In these regressions, we will use both indicator variables for year as well as a trend, so make a factor version of the `year` variable.
```{r deaths2}
deaths <- mutate(deaths, year_fct = factor(year))
```


## Table 5.2

Regression DD Estimates of MLDA-Induced Deaths among 18-20 year-olds, from 1970-1983

```{r dtypes}
dtypes <- c("all" = "All deaths", 
            "MVA" = "Motor vehicle accidents",
            "suicide" = "Suicide", 
            "internal" = "All internal causes")
```

Estimate the DD for MLDA for all causes of death in 18-20 year olds. 
Run the regression with `lm` and calculatate the cluster-robust standard errors
using `sandwich::vcovCL`.
Subset the data.
```{r data}
data <- filter(deaths, year <= 1983, agegr == "18-20 yrs", dtype == "all")
```
Run the OLS model.
```{r mod}
mod <- lm(mrate ~ 0 + legal + state + year_fct, 
          data = data)
```
Calculate cluster robuset coefficients.
These are calculcated using a different method than Stata's default, and thus will be slightly different than those reported in the book.
```{r vcov}
vcov <- vcovCR(mod, cluster = data[["state"]],
               type = "CR2")
coef_test(mod, vcov = vcov) %>%
  rownames_to_column(var = "term") %>%
  as_tibble() %>%
  select(term, estimate = beta, std.error = SE) %>%
  filter(term == "legal") %>%
  knitr::kable(digits = 2)
```

Function to calculate sclustered standard errors and return a tidy data frame of the coefficients and standard errors.
```{r cluster_se}
cluster_se <- function(mod, cluster, type = "CR2") {
  vcov <- vcovCR(mod, cluster = cluster, type = "CR2")
  coef_test(mod, vcov = vcov) %>%
    rownames_to_column(var = "term") %>%
    as_tibble() %>%
    select(term, estimate = beta, std.error = SE)
}
```


```{r run_mlda_dd}
run_mlda_dd <- function(i) {
  data <- filter(deaths, year <= 1983, agegr == "18-20 yrs", dtype == i) # nolint
  mods <- tribble(
    ~ name, ~ model,
    "No trends, no weights", lm(mrate ~ 0 + legal + state + year_fct, data = data),
    "Time trends, no weights",
      lm(mrate ~ 0 + legal + year_fct + state + state:year, data = data),
    "No trends, weights",
      lm(mrate ~ 0 + legal + year_fct + state, data = data, weights = pop),
    # "Time trends, weights",
    #   lm(mrate ~ 0 + legal + year_fct + state + state:year, data = data, weights = pop)
  ) %>%
    mutate(coefs = map(model, ~ cluster_se(.x, cluster = data[["state"]], type = "CR2"))) %>%
    unnest(coefs) %>%
    filter(term == "legal") %>%
    mutate(response = i) %>%
    select(name, response, estimate, std.error)
}
```

```{r mlda_dd}
mlda_dd <- map_df(names(dtypes), run_mlda_dd)
```

```{r mlda_dd_table}
mlda_dd %>%
  knitr::kable(digits = 2)
```


## Table 5.3

Regression DD Estimates of MLDA-Induced Deaths among 18-20 year-olds, from 1970-1983, controlling for Beer Taxes.
This is the analysis presented in @AngristPischke2014 Table 5.3.

```{r run_beertax}
run_beertax <- function(i) {
  data <- filter(deaths, year <= 1983, agegr == "18-20 yrs", 
                 dtype == i, !is.na(beertaxa))
  out <- tribble(
    ~ name, ~ model,
    "No time trends",
      lm(mrate ~ 0 + legal + beertaxa + year_fct + state, data = data),
    "Time trends",
      lm(mrate ~ 0 + legal + beertaxa + year_fct + state + state:year, data = data)
  ) %>%
    # calc culstered standard errors
    mutate(coefs = map(model, ~ cluster_se(.x, data[["state"]]))) %>%
    unnest(coefs) %>%
    filter(term %in% c("legal", "beertaxa")) %>%
    mutate(response = i) %>%    
    select(response, name, term, estimate, std.error)
}
```

```{r beertax}
beertax <- map_df(names(dtypes), run_beertax)
```

```{r tbl_beertax}
beertax %>%
  knitr::kable(digits = 2)
```

*Note:* I had trouble getting `sandwich::vcovCL` to estimate clustered standard errors for this regression.

## References

- <http://masteringmetrics.com/wp-content/uploads/2015/01/analysis.do>
- <http://masteringmetrics.com/wp-content/uploads/2015/01/ReadMe_MLDA_DD.txt>
