---
title: "Table 1.1""
author: "Jeffrey Arnold"
date: "3/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This reproduces the analyses in Table 1.1.
That table compares people with and without health insurance in the 2009 National Health Interview Survey (NHIS).

The code is derived from  [NHIS2009_hicompare.do](http://masteringmetrics.com/wp-content/uploads/2015/01/NHIS2009_hicompare.do).

First, load the prerequisite packages.
```{r message=FALSE}
library("tidyverse")
library("magrittr")
library("haven")
```

Load the data, and adjust a few of the columns to account for differences in 
how STATA and R store data.
```{r}
NHIS2009_clean <-  
  read_dta("data/NHIS/Data/NHIS2009_clean.dta") %>%
  mutate(age = as.numeric(age),
         serial = as.integer(serial),
         fml = as.logical(fml),
         marradult = as.logical(marradult))
```
Remove missing values.
```{r}
NHIS2009_clean <- NHIS2009_clean %>%
  filter(marradult, perweight != 0) %>%
  group_by(serial) %>%
  mutate(hi_hsb = mean(hi_hsb1, na.rm = TRUE)) %>%
  filter(!is.na(hi_hsb), !is.na(hi)) %>%
  mutate(female = sum(fml)) %>%
  filter(female == 1) %>%
  select(-female)
```

For the sample only include those between 26 and 59, and that are married, and remove single person households.
```{r}
NHIS2009_clean <- NHIS2009_clean %>%
  filter(between(age, 26, 59),
         marradult, adltempl >= 1)
         
```
Keep only single family households
```{r}
NHIS2009_clean %<>%
  group_by(serial) %>%
  filter(length(serial) > 1L) %>%
  ungroup()
```

Tables of wives and husbands by HI status. 
The weighting matches the original in Stata.
```{r}
NHIS2009_clean %>%
  group_by(fml) %>%
  # normalize person weights to match number of observations in each
  # group
  mutate(perweight = perweight / sum(perweight) * n()) %>%
  group_by(fml, hi) %>%
  summarise(n_wt = sum(perweight)) %>%
  group_by(fml) %>%
  mutate(prop = n_wt / sum(n_wt))

```

Comparison of stample statistics
```{r}
varlist <- c("hlth", "nwhite", "age", "yedu", "famsize", "empl", "inc")
NHIS2009_clean %>%
  # rlang::set_attrs with NULL removes attributes from columns.
  # this avoids a warning from gather about differing attributes
  map_dfc(~ rlang::set_attrs(.x, NULL)) %>%
  select(fml, hi, one_of(varlist)) %>% 
  gather(variable, value, -fml, -hi) %>%
  group_by(fml, hi, variable) %>%
  summarise(mean = mean(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE)) %>%
  gather(stat, value, -fml, -hi, -variable) %>%
  unite(stat_hi, stat, hi) %>%
  spread(stat_hi, value) %>%
  mutate(diff = mean_1 - mean_0)
```
