```{r include=FALSE, cache=FALSE}
set.seed(1014)
options(digits = 3)

# .nojekyll <- here::here("docs", ".nojekyll")
# if (!file.exists(.nojekyll)) {
#   close(open(file(.nojekyll, open = "w")))
# }

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE,
  autodep = TRUE,
  out.width = "70%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold"
)

options(dplyr.print_min = 6, dplyr.print_max = 6)

is_html <- knitr::opts_knit$get("rmarkdown.pandoc.to") == "html"
```
---
output: html_document
editor_options: 
  chunk_output_type: console
---
# RAND Health Insurance Experiment (HIE)

This provides code replicates the Tables 1.3 and 1.4 of @AngristPischke2014 which replicate the analyses from the RAND Health Insurance Experiment [@BrookWareEtAl1983,@Aron-DineEinavEtAl2013].

Load necessary libraries.
```{r libraries,message=FALSE}
library("tidyverse")
library("broom")
library("haven")
library("rlang")
library("clubSandwich")
```


## Table 1.3

@AngristPischke2014 Table 1.3 presents demographic and baseline health characteristics for subjects of the RAND Health Insurance Experiment (HIE).

Load the `rand` data.
```{r rand_sample}
data("rand_sample", package = "masteringmetrics")
```


Calculate the number in each plan:
```{r plantypes}
plantypes <- count(rand_sample, plantype) 
```
```{r results='asis'}
knitr::kable(plantypes)
```


For each variable variables, estimate the the difference in means between heath insurance plan types.
```{r}
varlist <- c("female", "blackhisp", "age", "educper", 
             "income1cpi", "hosp", "ghindx", "cholest", "diastol",
             "systol", "mhi", "ghindxx", 
             "cholestx", "diastolx", "systolx", "mhix")
		
```

Create column (1) with the mean and standard deviation of the "Catastrophic" plan,
```{r catastrophic_stats}
catastrophic_stats <- rand_sample %>%
  filter(plantype == "Catastrophic") %>%
  select(one_of(varlist)) %>%
  gather(variable, value) %>%
  group_by(variable) %>%
  summarise(Mean = mean(value, na.rm = TRUE), 
            `Std. Dev.` = sd(value, na.rm = TRUE))
```

```{r catastrophic_stats_tab, results='asis'}
knitr::kable(catastrophic_stats, digits = 3)
```

The difference in means between plans and the catastophic plan.
```{r}
calc_diffs <- function(x) {
  # programmatically create the formula for lm
  f <- quo(!!sym(x) ~ plantype)

  mod <- lm(f, data = rand_sample)
  out <- coef_test(mod, "CR2", cluster = rand_sample$fam_identifier) %>% 
    as_tibble() %>%
    rownames_to_column(var = "term") %>%   
    select(term, estimate = beta, std.error = SE)  
  out[["response"]] <- x
  out
}
```
```{r plantype_diffs}
plantype_diffs <- map_dfr(varlist, calc_diffs) %>%
  select(response, term, estimate, std.error) %>%
  mutate(term = str_replace(term, "^plantype", ""))
```

Create a table similar to @AngristPischke2014 Table 1.3.
```{r}
fmt_num <- function(x) {
  prettyNum(x, digits = 3, format = "f", big.mark = ",", drop0trailing = FALSE)
}

plantype_diffs %>% 
  mutate(estimate = str_c(fmt_num(estimate), " (", fmt_num(std.error), ")")) %>%
  select(-std.error) %>%
  spread(term, estimate) %>%
  knitr::kable(digits = 3)

```

Plot the difference-in-means of each plantype vs. catastrophic insurance.
```{r}
ggplot(filter(plantype_diffs, term != "(Intercept)"), 
              aes(x = term, y = estimate, 
                  ymin = estimate - 2 * std.error,
                  ymax = estimate + 2 * std.error)) +
  geom_hline(yintercept = 0, colour = "white", size = 1) +
  geom_pointrange() +  
  facet_grid(response ~ . , scales = "free_y")
  
```


## Table 1.4

Replicate @AngristPischke2014 Table 1.4 which presents health outcome and health expenditure results from the RAND HIE.

```{r rand_person_spend}
data("rand_person_spend", package = "masteringmetrics")
```

Correlate year variable from annual expenditures data to correct calendar year in order to adjust for inflation.
```{r person_spend_expyear}
rand_person_spend <- 
  mutate(rand_person_spend,
         expyear = indv_start_year + year - 1)
```

Adjust spending for inflation.
The CPI adjustment values below are based on the June CPI from 1991 (see table found at http://www.seattle.gov/financedepartment/cpi/historical.htm ).

```{r cpi}
cpi <- tribble(
  ~ year, ~ cpi,
  1973, 3.07,
  1974, 2.76,
  1975, 2.53,
  1976, 2.39,
  1977, 2.24,
  1978, 2.09,
  1979, 1.88,
  1980, 1.65,
  1981, 1.5,
  1982, 1.41,
  1983, 1.37,
  1984, 1.31,
  1985, 1.27
)
```
```{r person_spend_cpi}
rand_person_spend <- 
  left_join(rand_person_spend, 
            cpi, by = c("expyear" = "year")) %>%
  mutate(out_inf = outsum * cpi,
         inpdol_inf = inpdol * cpi)
```

Add a total spending variable.
```{r person_spend_tot_inf}
rand_person_spend <- mutate(rand_person_spend,
                       tot_inf = inpdol_inf + out_inf)
```
Add a variable for any health insurance (free, Individual deductible, or cost-sharing):
```{r person_spend}
rand_person_spend <- mutate(rand_person_spend, 
                       any_ins = plantype != "Catastrophic")
```

Count the number of observations in each plan-type,
```{r}
count(rand_person_spend, plantype)
```
and any-insurance,
```{r}
count(rand_person_spend, any_ins)
```

Create a list of response variables.
```{r}
varlist <- c("ftf", "out_inf", "totadm", "inpdol_inf", "tot_inf")
```

Calculate the mean and standard deviation for those receiving catastrophic insurance.
```{r}
rand_person_spend %>%
  filter(plantype == "Catastrophic") %>%
  select(one_of(varlist)) %>%
  gather(response, value) %>%
  group_by(response) %>%
  summarise(Mean = mean(value, na.rm = TRUE),
            `Std. Dev.` = sd(value, na.rm = TRUE))
```

Calculate the difference in means between plans and the catastophic plan.
```{r}
calc_diffs <- function(x) {
  # programmatically create the formula
  f <- quo(!!sym(x) ~ plantype)

  mod <- lm(f, data = rand_person_spend)
  out <- coef_test(mod, "CR2", cluster = rand_person_spend$fam_identifier) %>% 
    as_tibble() %>%
    rownames_to_column(var = "term") %>%   
    select(term, estimate = beta, std.error = SE)
  out[["response"]] <- x
  out
}
```
```{r}
person_diffs <- map_dfr(varlist, calc_diffs) %>%
  select(response, term, estimate, std.error) %>%
  mutate(term = str_replace(term, "^plantype", ""))

```
Standard errors are clustered by family identifier using the **clubSandwich** package.

Print the table. If this were an actual publication, I'd make it nicer.
```{r}
fmt_num <- function(x) {
  prettyNum(x, digits = 3, format = "f", big.mark = ",", drop0trailing = FALSE)
}

person_diffs %>% 
  mutate(estimate = str_c(fmt_num(estimate), " (", fmt_num(std.error), ")")) %>%
  select(-std.error) %>%
  spread(term, estimate) %>%
  knitr::kable(digits = 3)

```

Additionally we could plot the difference-in-means of each plantype vs. catastrophic insurance.
```{r}
ggplot(filter(person_diffs, term != "(Intercept)"), 
              aes(x = term, y = estimate, 
                  ymin = estimate - 2 * std.error,
                  ymax = estimate + 2 * std.error)) +
  geom_hline(yintercept = 0, colour = "white", size = 1) +
  geom_pointrange() +  
  facet_grid(response ~ . , scales = "free_y")
  
```

## References {-}

- <https://www.icpsr.umich.edu/icpsrweb/NACDA/studies/6439/version/1>
- <http://masteringmetrics.com/wp-content/uploads/2015/01/ReadMe_RAND.txt>
- <http://masteringmetrics.com/wp-content/uploads/2015/01/Code.zip>
```{r include=FALSE, cache=FALSE}
```
