# Tables 1.3 and 1.4

```{r libraries,message=FALSE}
library("tidyverse")
library("broom")
library("haven")
```

```{r}
dta_variables <- function(x) {
  stata_attrs <- c("format.stata", "label")
  out <- map_dfr(x, function(x) {
    as_tibble(compact(attributes(x)[stata_attrs]))
  })
  out[["variable"]] <- names(x)
  out[["class"]] <- map(x, class)
  out <- select(out, variable, everything())
  out
}
```

## Table 1.3

```{r FILENAME}
FILENAME <- "data/Data/rand_initial_sample_2.dta"
```

```{r rand}
rand <- read_dta(FILENAME) %>%
  # delete Stata attributes
  map_dfc(~ rlang::set_attrs(.x, NULL))
```

The `plantype` variable takes four values,

------- ------------------
1       Free
2       Deductible
3       Coinsurance
4       Catastrophic
------- -------------------

Create a more interpretable version of plantype, and 
ensure that the free plan does not give 0's to anyone without a plan.
```{r rand_plantype}
rand <- rand %>%
  filter(!is.na(plantype)) %>%
  # plantype to a factor variable
  mutate(plantype = factor(plantype, 
                           labels = c("Free", "Deductible", 
                                      "Coinsurance", "Catastrophic"))) %>%
  # reorder so that "Catastrophic is first
  mutate(plantype = fct_relevel(plantype, 
                                "Catastrophic", "Deductible",
                                "Coinsurance", "Free")) %>%
  # indicator variable for any insurance
  mutate(any_ins = plantype != "Catastrophic")
```

Calculate the number in each plan:
```{r plantypes}
plantypes <- count(rand, plantype) 
```
```{r results='asis'}
knitr::kable(plantypes)
```


Create column with an integer family identifier.
```{r rand_famid}
rand <- mutate(rand,
               famid = str_sub(fam_identifier, 3) %>% 
                 str_replace("[A-Z]", "") %>%
                 as.integer())
```


For variables, difference in means between plantypes
```{r}
varlist <- c("female", "blackhisp", "age", "educper", "income1cpi", "hosp", 
             "ghindx", "cholest", "diastol", "systol", "mhi", "ghindxx", 
             "cholestx", "diastolx", "systolx", "mhix")
		
```

Column (1) with the mean and standard deviation of the "Catastrophic" plan,
```{r catastrophic_stats}
catastrophic_stats <- rand %>%
  filter(plantype == "Catastrophic") %>%
  select(one_of(varlist)) %>%
  gather(variable, value) %>%
  group_by(variable) %>%
  summarise(Mean = mean(value, na.rm = TRUE), 
            `Std. Dev.` = sd(value, na.rm = TRUE))
```

```{r catastrophic_stats_tab, results='asis'}
knitr::kable(catastrophic_stats, digits = 3)

```


The difference in means between plans and the catastophic plan.
```{r}
calc_diffs <- function(x) {
  # programmatically create the formula
  f <- quo_squash(quo(!!sym(x) ~ plantype))
  # this would also work
  # f <- as.formula(str_c(x, " ~ plantype_1 + plantype_2 + plantype_3"))

  mod <- lm(f, data = rand)
  out <- tidy(mod)
  out[["response"]] <- x
  out
}

x_diffs <- map_dfr(varlist, calc_diffs) %>%
  select(response, term, estimate, std.error) %>%
  mutate(term = str_replace(term, "^plantype", ""))

```
Note: the above code currently does not cluster by family ID as in the original analysis
so the standard errors will be different.

Print the table. This could be done much nicer, ... but for now.
```{r}
fmt_num <- function(x) {
  prettyNum(x, digits = 3, format = "f", big.mark = ",", drop0trailing = FALSE)
}

x_diffs %>% 
  mutate(estimate = str_c(fmt_num(estimate), " (", fmt_num(std.error), ")")) %>%
  select(-std.error) %>%
  spread(term, estimate) %>%
  knitr::kable(digits = 3)

```

Plot the difference-in-means of each plantype vs. catastrophic insurance.
```{r plot_plantype_diffs}
ggplot(filter(x_diffs, term != "(Intercept)"), 
              aes(x = term, y = estimate, 
                  ymin = estimate - 2 * std.error,
                  ymax = estimate + 2 * std.error)) +
  geom_hline(yintercept = 0, colour = "white", size = 1) +
  geom_pointrange() +  
  facet_grid(response ~ . , scales = "free_y")
  
```


## Table 1.4

Load person-year data
```{r person_years}
person_years <- read_dta(here::here("data/Data/person_years.dta")) %>%
    map_dfc(~ rlang::set_attrs(.x, NULL))
```
Load annual spending on hospital visits
```{r}
annual_spend <- read_dta(here::here("data/Data/annual_spend.dta")) %>%
  map_dfc(~ rlang::set_attrs(.x, NULL))
```

Inner join the person-year and annual-spending data on person identifiers and years,
```{r}
person_spend <- inner_join(person_years, annual_spend, 
                           by = c("person", "year"))
```

Group plans into four types

- 1-Free, 
- 2-Individual Deductible
- 3-Cost Sharing (25%/50%)
- 4-Catostrophic (Fam Deductible) (95%/100%)

```{r person_plantype}
person_spend <-
  mutate(person_spend,
         plantype = case_when(
          plan == 24 ~ "Free",
          plan %in% c(1, 5) ~ "Deductible",
          plan >= 2 & plan <= 4 ~ "Catastrophic",
          plan >= 6 & plan <= 8 ~ "Catastrophic",
          plan >= 9 & plan <= 23 ~ "Cost Sharing"
        )) %>%
  # reorder levels so Catastrophic is first
  mutate(plantype = fct_relevel(plantype, 
                                "Catastrophic",
                                "Deductible",
                                "Cost Sharing",
                                "Free"))
```

Attempt to correlate year variable from annual expenditures data to correct calendar year in order to adjust for inflation.
```{r}
person_spend <- mutate(person_spend,
                       expyear = indv_start_year + year - 1)
```

Adjust spending for inflation.
The CPI adjustment values below are based on the June CPI from 1991 (see table found at http://www.seattle.gov/financedepartment/cpi/historical.htm ).

```{r cpi}
cpi <- tribble(
  ~ year, ~ cpi,
  1973, 3.07,
  1974, 2.76,
  1975, 2.53,
  1976, 2.39,
  1977, 2.24,
  1978, 2.09,
  1979, 1.88,
  1980, 1.65,
  1981, 1.5,
  1982, 1.41,
  1983, 1.37,
  1984, 1.31,
  1985, 1.27
)
```
```{r}
person_spend <- 
  left_join(person_spend, 
            cpi, by = c("expyear" = "year")) %>%
  mutate(out_inf = outsum * cpi,
         inpdol_inf = inpdol * cpi)
```

Add a total spending variable.
```{r}
person_spend <- mutate(person_spend,
                       tot_inf = inpdol_inf + out_inf)
```
Add a variable for any health insurance (free, Individual deductible, or cost-sharing):
```{r}
person_spend <- mutate(person_spend, 
                       any_ins = plantype != "Catastrophic")
```


Count the number of observations in each plan-type,
```{r person_plantype}
count(person_spend, plantype)
```
and any-insurance,
```{r person_plantype}
count(person_spend, any_ins)
```

Create a list of response variables.
```{r}
varlist <- c("ftf", "out_inf", "totadm", "inpdol_inf", "tot_inf")
```

Mean and standard deviation for those receiving catastrophic insurance,
```{r}
person_spend %>%
  filter(plantype == "Catastrophic") %>%
  select(one_of(varlist)) %>%
  gather(response, value) %>%
  group_by(response) %>%
  summarise(Mean = mean(value, na.rm = TRUE),
            `Std. Dev.` = sd(value, na.rm = TRUE))
```

Calculate a difference in means for all
The difference in means between plans and the catastophic plan.
```{r}
calc_diffs <- function(x) {
  # programmatically create the formula
  f <- quo_squash(quo(!!sym(x) ~ plantype))
  # this would also work
  # f <- as.formula(str_c(x, " ~ plantype_1 + plantype_2 + plantype_3"))

  mod <- lm(f, data = person_spend)
  out <- tidy(mod)
  out[["response"]] <- x
  out
}

person_diffs <- map_dfr(varlist, calc_diffs) %>%
  select(response, term, estimate, std.error) %>%
  mutate(term = str_replace(term, "^plantype", ""))

```
Note: the above code currently does not cluster by family ID as in the original analysis
so the standard errors will be different.

Print the table. This could be done much nicer, ... but for now.
```{r}
fmt_num <- function(x) {
  prettyNum(x, digits = 3, format = "f", big.mark = ",", drop0trailing = FALSE)
}

person_diffs %>% 
  mutate(estimate = str_c(fmt_num(estimate), " (", fmt_num(std.error), ")")) %>%
  select(-std.error) %>%
  spread(term, estimate) %>%
  knitr::kable(digits = 3)

```

Plot the difference-in-means of each plantype vs. catastrophic insurance.
```{r plot_plantype_diffs}
ggplot(filter(person_diffs, term != "(Intercept)"), 
              aes(x = term, y = estimate, 
                  ymin = estimate - 2 * std.error,
                  ymax = estimate + 2 * std.error)) +
  geom_hline(yintercept = 0, colour = "white", size = 1) +
  geom_pointrange() +  
  facet_grid(response ~ . , scales = "free_y")
  
```
