---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Tables 1.3 and 1.4

Load necessary libraries.
```{r libraries,message=FALSE}
library("tidyverse")
library("broom")
library("haven")
library("rlang")
```

Calculate the number in each plan:
```{r plantypes}
plantypes <- count(rand, plantype) 
```
```{r results='asis'}
knitr::kable(plantypes)
```


For variables, difference in means between plantypes
```{r}
varlist <- c("female", "blackhisp", "age", "educper", "income1cpi", "hosp", 
             "ghindx", "cholest", "diastol", "systol", "mhi", "ghindxx", 
             "cholestx", "diastolx", "systolx", "mhix")
		
```

Create column (1) with the mean and standard deviation of the "Catastrophic" plan,
```{r catastrophic_stats}
catastrophic_stats <- rand %>%
  filter(plantype == "Catastrophic") %>%
  select(one_of(varlist)) %>%
  gather(variable, value) %>%
  group_by(variable) %>%
  summarise(Mean = mean(value, na.rm = TRUE), 
            `Std. Dev.` = sd(value, na.rm = TRUE))
```

```{r catastrophic_stats_tab, results='asis'}
knitr::kable(catastrophic_stats, digits = 3)

```


The difference in means between plans and the catastophic plan.
```{r}
calc_diffs <- function(x) {
  # programmatically create the formula
  f <- quo(!!sym(x) ~ plantype)
  # this would also work
  # f <- as.formula(str_c(x, " ~ plantype_1 + plantype_2 + plantype_3"))

  mod <- lm(f, data = rand)
  out <- tidy(mod)
  out[["response"]] <- x
  out
}

x_diffs <- map_dfr(varlist, calc_diffs) %>%
  select(response, term, estimate, std.error) %>%
  mutate(term = str_replace(term, "^plantype", ""))

```
Note: the above code currently does not cluster by family ID as in the original analysis
so the standard errors will be different.

Print the table. This could be done much nicer, ... but for now.
```{r}
fmt_num <- function(x) {
  prettyNum(x, digits = 3, format = "f", big.mark = ",", drop0trailing = FALSE)
}

x_diffs %>% 
  mutate(estimate = str_c(fmt_num(estimate), " (", fmt_num(std.error), ")")) %>%
  select(-std.error) %>%
  spread(term, estimate) %>%
  knitr::kable(digits = 3)

```

Plot the difference-in-means of each plantype vs. catastrophic insurance.
```{r}
ggplot(filter(x_diffs, term != "(Intercept)"), 
              aes(x = term, y = estimate, 
                  ymin = estimate - 2 * std.error,
                  ymax = estimate + 2 * std.error)) +
  geom_hline(yintercept = 0, colour = "white", size = 1) +
  geom_pointrange() +  
  facet_grid(response ~ . , scales = "free_y")
  
```


## Table 1.4

```{r}
data("rand_person_spend", package = "masteringmetrics")
```

 correlate year variable from annual expenditures data to correct calendar year in order to adjust for inflation.
```{r}
person_spend <- mutate(person_spend,
                       expyear = indv_start_year + year - 1)
```

Adjust spending for inflation.
The CPI adjustment values below are based on the June CPI from 1991 (see table found at http://www.seattle.gov/financedepartment/cpi/historical.htm ).

```{r cpi}
cpi <- tribble(
  ~ year, ~ cpi,
  1973, 3.07,
  1974, 2.76,
  1975, 2.53,
  1976, 2.39,
  1977, 2.24,
  1978, 2.09,
  1979, 1.88,
  1980, 1.65,
  1981, 1.5,
  1982, 1.41,
  1983, 1.37,
  1984, 1.31,
  1985, 1.27
)
```
```{r}
person_spend <- 
  left_join(person_spend, 
            cpi, by = c("expyear" = "year")) %>%
  mutate(out_inf = outsum * cpi,
         inpdol_inf = inpdol * cpi)
```

Add a total spending variable.
```{r}
person_spend <- mutate(person_spend,
                       tot_inf = inpdol_inf + out_inf)
```
Add a variable for any health insurance (free, Individual deductible, or cost-sharing):
```{r}
person_spend <- mutate(person_spend, 
                       any_ins = plantype != "Catastrophic")
```

Count the number of observations in each plan-type,
```{r}
count(person_spend, plantype)
```
and any-insurance,
```{r}
count(person_spend, any_ins)
```

Create a list of response variables.
```{r}
varlist <- c("ftf", "out_inf", "totadm", "inpdol_inf", "tot_inf")
```

Mean and standard deviation for those receiving catastrophic insurance,
```{r}
person_spend %>%
  filter(plantype == "Catastrophic") %>%
  select(one_of(varlist)) %>%
  gather(response, value) %>%
  group_by(response) %>%
  summarise(Mean = mean(value, na.rm = TRUE),
            `Std. Dev.` = sd(value, na.rm = TRUE))
```

Calculate the difference in means between plans and the catastophic plan.
```{r}
calc_diffs <- function(x) {
  # programmatically create the formula
  f <- quo_squash(quo(!!sym(x) ~ plantype))
  # this would also work
  # f <- as.formula(str_c(x, " ~ plantype_1 + plantype_2 + plantype_3"))

  mod <- lm(f, data = person_spend)
  out <- tidy(mod)
  out[["response"]] <- x
  out
}

person_diffs <- map_dfr(varlist, calc_diffs) %>%
  select(response, term, estimate, std.error) %>%
  mutate(term = str_replace(term, "^plantype", ""))

```
Note: the above code currently does not cluster by family identifier as in the original analysis
so the standard errors will be different.

Print the table. If this were an actual publication, I'd make it nicer.
```{r}
fmt_num <- function(x) {
  prettyNum(x, digits = 3, format = "f", big.mark = ",", drop0trailing = FALSE)
}

person_diffs %>% 
  mutate(estimate = str_c(fmt_num(estimate), " (", fmt_num(std.error), ")")) %>%
  select(-std.error) %>%
  spread(term, estimate) %>%
  knitr::kable(digits = 3)

```

Additionally we could plot the difference-in-means of each plantype vs. catastrophic insurance.
```{r}
ggplot(filter(person_diffs, term != "(Intercept)"), 
              aes(x = term, y = estimate, 
                  ymin = estimate - 2 * std.error,
                  ymax = estimate + 2 * std.error)) +
  geom_hline(yintercept = 0, colour = "white", size = 1) +
  geom_pointrange() +  
  facet_grid(response ~ . , scales = "free_y")
  
```
