---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Twins and Returns to Schooling

Estimates of the returns to schooling for Twinsburg twins [@AshenfelterKrueger1994;@AshenfelterRouse1998].
This replicates the analysis in Table 6.2 of *Mastering 'Metrics*.

```{r setup,message=FALSE}
library("tidyverse")
library("sandwich")
library("lmtest")
library("AER")
```

Load `twins` data.
```{r pubtwins}
data("pubtwins", package = "masteringmetrics")
```

Run a regression of log wage on controls (Column 1 of Table 6.2).
```{r mod1}
mod1 <- lm(lwage ~ educ + poly(age, 2) + female + white, data = pubtwins)
coeftest(mod1, vcov = vcovHC(mod1))
```
*Note:* The `age` coefficients are different (but equivalent) to those reported in the Table due to the the use of `poly(age, .)`, which calculates orthogonal polynomials.

Run regression of the difference in log wage between twins on the difference in education (Column 2 of Table 6.2).
```{r mod2}
mod2 <- lm(dlwage ~ deduc, data = filter(pubtwins, first == 1))
coeftest(mod2, vcov = vcovHC(mod2))
```

Run a regression of log wage on controls, instrumenting education with twin's education (Column 3 of Table 6.2).
```{r mod3}
mod3 <- ivreg(lwage ~ educ + poly(age, 2) + female + white | 
                . - educ + educt, data = pubtwins)
coeftest(mod3, vcov = vcovHC(mod3))
```
*Note:* The coefficient for years of education is slightly different than that reported in the book.

Run a regression of the difference in wage, instrumenting the difference in years of education with twin's education (Column 4 of Table 6.2).
```{r mod4}
mod4 <- ivreg(dlwage ~ deduc | deduct,
              data = filter(pubtwins, first == 1))
coeftest(mod4, vcov = vcovHC(mod4))
```
*Note:* The coefficient for years of education is slightly different than that reported in the book.

## References {-}

-   <http://masteringmetrics.com/wp-content/uploads/2015/02/ReadMe_Twinsburg.txt>
-   <http://masteringmetrics.com/wp-content/uploads/2015/02/twins.do>

